---
title: 'Activity: Gene Expression'
author: "Maggie Linn"
output:
  html_document:
    df_print: paged
---

```{r, message = F, warning = F}
# include important libraries 
library(mosaic)
library(tidyverse)
library(lubridate)
library(dplyr)
library(dcData)
```

```{r}
data("NCI60")
Narrow <-
  NCI60 %>%
  tidyr::pivot_longer(cols = -Probe, names_to = "cellLine", values_to = "expression")

CellTypes <-
  NCI60cells %>%
  select(cellLine, tissue) %>%
  mutate(cellLine = gsub(pattern = "\\:", replacement = ".", x = as.character(cellLine)))

Narrow <- 
  Narrow %>%
  inner_join(CellTypes)

Probe_TOP3A <-
  Narrow %>%
  filter(Probe == "TOP3A")

SummaryStats <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn_expr = exp(mean(expression, na.rm = TRUE)))
```


## Recreate Figure 18.1 

```{r}
SummaryStats %>% 
  ggplot(aes(x = tissue, y = mn_expr)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```


##  Recreate Figure 18.2

```{r}
Probe_TOP3A %>% 
ggplot(aes(x = tissue, y = exp(expression))) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```


## Critique of Figure 18.1



## Recreate Figure 18.3

```{r}
SummaryStats %>% 
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_bar(stat = "identity", fill = "gray", color = NA) +
  geom_point(data = Probe_TOP3A, aes(x = tissue, y = exp(expression))) + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

##  Recreate Figure 18.4

```{r}
SummaryStats <-  
      SummaryStats %>%  
      mutate(tissue = reorder(tissue, mn_expr)) 
SummaryStats <-
  Probe_TOP3A %>%
  group_by(tissue) %>%
  summarise(mn_expr = mean(expression,  na.rm = TRUE),
            se = sd(expression, na.rm = TRUE) / sqrt(n())) 
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>%
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_bar(stat = "identity", alpha = 0.2) +
  geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

##  Recreate Figure 18.5

```{r}
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>% 
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = Probe_TOP3A, aes(x = tissue, y = exp(expression))) + 
    geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

```


## Your turn: Pick your own probe and make a figure like that of Figure 18.5.

```{r}
Probe_ZZZ3 <-
  Narrow %>%
  filter(Probe == "ZZZ3")
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>% 
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = Probe_ZZZ3, aes(x = tissue, y = exp(expression))) + 
    geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

```

## Recreate figure 18.6 

```{r}
#needed uses-defined function to complete this problem 
r2 <- function(data) {
  mosaic::rsquared(lm(data$expression ~ data$tissue))
}
ProbeR2 <- 
  Narrow %>% 
  group_by(Probe) %>%
  dplyr::do(probe_rsq = r2(.)) %>%
  mutate(r2 = unlist(probe_rsq))
Actual <-
  ProbeR2 %>%
  arrange(desc(r2)) %>%
  head(30) 
Actual %>%
  ggplot(aes(x = reorder(Probe, desc(r2)), y = r2)) +
  geom_point() + 
  xlab("Probe") + 
  ylab("R-squared") + 
  ggtitle("Probes with large R-squared for expression level explained by tissue type") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Your turn: Choose one of the probes with high $R^2$. Plot out expression versus tissue type, just as Figure 18.5 shows it for TOP3A

```{r}
ProbeR2 <- 
  Narrow %>% 
  group_by(Probe) %>%
  dplyr::do(probe_rsq = r2(.)) %>%
  mutate(r2 = unlist(probe_rsq))
Probe_CD53 <-
  Narrow %>%
  filter(Probe == "CD53")
SummaryStats <- 
  SummaryStats %>%
  mutate(top = mn_expr + 2 * se,
         bottom = mn_expr - 2 * se) 
SummaryStats %>% 
  ggplot(aes(x = tissue, y = exp(mn_expr))) +
  geom_point(data = Probe_CD53, aes(x = tissue, y = exp(expression))) + 
    geom_errorbar(aes(x = tissue, 
                    ymax = exp(top), 
                    ymin = exp(bottom)), width = 0.5) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

```


#### Do you see a qualitative difference between the graph of your high $R^2$ probe and Figure 18.5?

There is a qualitative difference between the graphs because the error bar is much higher on figure 18.5 than the one I created with the probe with a high R^2 value.

## Recreate Figure 18.7 

(make sure to have **2** densities on the **same** plot)

```{r}

NullR2 <-
  Narrow %>%
  group_by(Probe) %>%
  mutate(expression = mosaic::shuffle(expression)) %>%
  group_by(Probe) %>%
  do(r2 = r2(.)) %>%
  mutate(r2 = unlist(r2))
ProbeR2 %>%
  ggplot(aes(x = r2)) + 
  geom_density(fill = "gray30", color = NA) +
  geom_density(data = NullR2, aes(x = r2), 
               fill = "gray80", alpha = .75, color = NA)

```

## Recreate Figure 18.8


```{r}

Null <-
  NullR2 %>% 
  arrange(desc(r2)) %>%
  head(30) 

# append the 30 highest `Null` values to the `Actual` data
Actual$null <- Null$r2

Actual %>%
  ggplot(aes(x = reorder(Probe, desc(r2)), y = r2)) +
  geom_point() + 
  geom_point(aes(y = null), color = "gray50") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



