---
title: "Activity: Gene Expression"
author: "YOUR NAME HERE"
output: html_notebook
---

```{r, message = F, warning = F}
# include important libraries 

```



## Recreate Figure 18.1 

```{r}


```


##  Recreate Figure 18.2

```{r}


```


## Critique of Figure 18.1

My answer here...... 


## Recreate Figure 18.3

```{r}


```

##  Recreate Figure 18.4

```{r}


```

##  Recreate Figure 18.5

```{r}


```


## Your turn: Pick your own probe and make a figure like that of Figure 18.5.

```{r}


```

## Recreate figure 18.6 

```{r}
#needed uses-defined function to complete this problem 
r2 <- function(data) {
  mosaic::rsquared(lm(data$expression ~ data$tissue))
}




```

## Your turn: Choose one of the probes with high $R^2$. Plot out expression versus tissue type, just as Figure 18.5 shows it for TOP3A

```{r}


```


#### Do you see a qualitative difference between the graph of your high $R^2$ probe and Figure 18.5?

My answer here.... 


## Recreate Figure 18.7 

(make sure to have **2** densities on the **same** plot)

```{r}



```

## Recreate Figure 18.8


```{r}



```



